---
title: "Datafest markdown(TAEHEE)"
author: "TAEHEE KIM", "Nicole", "Samantha Alejandre"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## client data
```{r}
library(readxl)
rm(list = ls())
clients <- read.csv("clients.csv")
head(clients)
```

```{r}
# Load required libraries
library(tidyverse)
library(ggplot2)
library(ggthemes)
library(maps)
# Clean the dataset to only include relevant columns
client_locs <- clients %>% 
  select(ClientUno, StateAbbr, County, PostalCode, AnnualIncome)
# Convert PostalCode to character and pad with leading zeros to ensure 5 digits
client_locs$PostalCode <- str_pad(as.character(client_locs$PostalCode), width = 5, side = "left", pad = "0")

client_locs<- na.omit(client_locs)
client_locs<- client_locs[!client_locs$AnnualIncome == "NULL", ] 
client_locs

# Merge with map data to get state names and coordinates
us_map <- map_data("state")
client_locs <- client_locs %>%
  left_join(us_map, by = c("StateAbbr"= "region"))

us_map 
# Group by state and count number of clients
state_counts <- client_locs %>%
  group_by(StateAbbr) %>%
  summarise(num_clients = n()) %>%
  arrange(desc(num_clients))

state_counts

# Plot the data on a map



```
```{r}
#install.packages("usmap")
library(usmap) #import the package
library(ggplot2) #use ggplot2 to add layer 

client_in <- client_locs %>% 
  select(StateAbbr, AnnualIncome)
client_in<- na.omit(client_in)

# create a sample dataset
income <- data.frame(StateAbbr = client_in$StateAbbr, 
                 AnnualIncome = client_in$AnnualIncome)

# calculate the average hours per state
#mean vs median, outliers (mention )
income_avg <- aggregate(as.numeric(AnnualIncome) ~ StateAbbr, data = income, FUN = median)
colnames(income_avg)= c("state", "AverageIncome")
colnames(state_counts)= c("state", "num_clients")
# create the bar chart
income
income_avg
state_counts
# Join the two data frames
income_data <- left_join(state_counts, income_avg, by = c("state"= "state"))

# Calculate the attorney-to-client ratio
income_data$income_client_ratio <-  income_data$AverageIncome/ income_data$num_clients

income_ratio <- data.frame(StateAbbr = income_data$state,income_data$income_client_ratio )

income_ratio

ggplot(income_ratio, aes(x = StateAbbr, y = income_data$income_client_ratio)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Ratio") +
  coord_flip()


```

```{r}

library(usmap)
library(ggplot2)
colnames(state_counts)= c("state", "num_clients")
plot_usmap(data = state_counts, values = "num_clients", include = state_counts$state , color = "blue") + scale_fill_continuous(low = "white", high = "blue", label = scales::comma) + theme(legend.position = "right")
```

```{r}
attorney<- read_csv("attorneys.csv")
head(attorney)

attorney_locs<- attorney %>% select(AttorneyUno, StateAbbr, County, PostalCode) 

attorney_locs$PostalCode <- str_pad(as.character(attorney_locs$PostalCode), width = 5, side = "left", pad = "0")

attorney_locs<- na.omit(attorney_locs)

state_counts3 <- attorney_locs %>%
  group_by(StateAbbr) %>%
  summarise(num_attorney = n()) %>%
  arrange(desc(num_attorney))

state_counts3

```
```{r}
colnames(state_counts3)= c("state", "num_attorney")
plot_usmap(data = state_counts3, values = "num_attorney", include = state_counts3$state , color = "blue") + scale_fill_continuous(low = "white", high = "blue", label = scales::comma) + theme(legend.position = "right")
```
```{r}
colnames(state_ratio)= c("state", "ClientToAttorney")
state_ratio
p1<- plot_usmap(data = state_ratio, values = "ClientToAttorney", include = state_ratio$state , color = "blue") + scale_fill_continuous(low = "white", high = "blue", label = scales::comma) + theme(legend.position = "right")
```
```{r}
p2 <- plot_usmap(data = income_data, values = "AverageIncome", include = income_data$state , color = "red") + scale_fill_continuous(low = "white", high = "red", label = scales::comma) + theme(legend.position = "right")
```

```{r}
# Combine the ggplot objects using ggarrange()
install.packages("ggpubr")
library(ggpubr)
ggarrange(p1, p2, ncol = 2, nrow = 2)
```



```{r}
attorney_time<- read_csv("attorneytimeentries.csv")
head(attorney_time)
attorney_time<- attorney_time %>% select( StateAbbr, Hours, EnteredOnUtc) 
attorney_time
#number attonery per client request
#external data population centsus per state, num of attery per captial 
```

```{r}
library(ggplot2)

# create a sample dataset
df <- data.frame(StateAbbr = attorney_time$StateAbbr, 
                 Hours = attorney_time$Hours)

# calculate the average hours per state
df_avg <- aggregate(Hours ~ StateAbbr, data = df, FUN = mean)

df_avg[nrow(df_avg) + 1,] <- list("ID", 0)

# create the bar chart
ggplot(df_avg, aes(x = StateAbbr, y = Hours)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Average Hours") +
  coord_flip()

```


```{r}
library(ggplot2)



# create a sample dataset
df <- data.frame(StateAbbr = attorney_time$StateAbbr, 
                 Hours = attorney_time$Hours)

# calculate the average hours per state
df_avg <- aggregate(Hours ~ StateAbbr, data = df, FUN = mean)
colnames(df_avg)= c("state", "average_hours")


# create the bar chart
ggplot(df_avg, aes(x = StateAbbr, y = Hours)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Average Hours") +
  coord_flip()

```
```{r}
beforecovid<- read_csv("G:/My Drive/DataFest/Shared Work/New subdata(Eric. Before covid, after covid)/
                       extractbeforecovid(attorney).csv")
bef_covid <- data.frame(StateAbbr = beforecovid$StateAbbr, 
                 Hours = beforecovid$Hours)

bef_covid
before_avg <- aggregate(Hours ~ StateAbbr, data = bef_covid, FUN = mean)
ggplot(before_avg, aes(x = StateAbbr, y = Hours)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Average Hours") +
  coord_flip()
```

```{r}
aftercovid<- read_csv("extractaftercovid(attorney).csv")
after_cov<- data.frame(StateAbbr = aftercovid$StateAbbr, 
                 Hours = aftercovid$Hours)
after_avg <- aggregate(Hours ~ StateAbbr, data = after_cov, FUN = mean)
ggplot(after_avg, aes(x = StateAbbr, y = Hours)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Average Hours") +
  coord_flip()
library(yarrr)
hgA <- hist(after_avg$Hours, breaks = 12)
hgB <- hist(before_avg$Hours)
range(c(hgA$breaks, hgB$breaks))
max(c(hgA$count, hgB$count))
plot(hgA, col = yarrr::transparent("red", trans.val = .9), xlim = c(0.1, 1.4), ylim = c(0,18), fill= "transparent")
plot(hgB, add = TRUE, col = yarrr::transparent("blue", trans.val = .9), fill= "transparent")
```
```{r}
ggplot(hgB, aes(f0)) + 
    geom_histogram(data = after_avg, fill = "red", alpha = 0.2) + 
    geom_histogram(data = before_avg, fill = "blue", alpha = 0.2) 
    
```

beforeaverage<- read_csv("before_avg.csv")
View(beforeaverage)


afteraverage<- read_csv("after_avg.csv")
View(afteraverage)

afterminusbefore_avg <-data.frame( StateAbbr= afteraverage$StateAbbr, DIfference= afteraverage$Hours - beforeaverage$Hours)

View(afterminusbefore_avg)



#after-before ggplot(don't erase this!)
after-before_avg <- aggregate(Hours ~ StateAbbr, data = bef_covid, FUN = mean) - aggregate(Hours ~ StateAbbr, data = after_cov, FUN = mean)
ggplot(after-before_avg, aes(x = StateAbbr, y = Hours)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Average Hours") +
  coord_flip()

write.csv(before_avg, "before_avg.csv", row.names=FALSE, quote=FALSE)
write.csv(after_avg, "after_avg.csv", row.names=FALSE, quote=FALSE)

ggplot(afterminusbefore_avg, aes(x = StateAbbr, y = DIfference)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Difference") +
  coord_flip()
  
```{r}
ggplot(afterminusbefore_avg, aes(x = StateAbbr, y = DIfference)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Difference") +
  coord_flip()
```
```{r}
ggplot(afterminusbefore_avg, aes(x = StateAbbr, y = DIfference, fill = DIfference)) +
  geom_bar(stat = "identity", width = 0.9) +
  labs(x = "State", y = "Difference") +
  scale_fill_gradient(low = "white", high = "darkblue") +
  coord_flip() +
  theme_minimal() + 
  theme(axis.title.x = element_text(size = 12, face = "bold", color = "black"),
        axis.title.y = element_text(size = 12, face = "bold", color = "black"))

```




```{r}
# Join the two data frames
state_data <- left_join(state_counts, state_counts3, by = c("state" = "state"))

# Calculate the attorney-to-client ratio
state_data$attorney_client_ratio <-  state_data$num_clients/ state_data$num_attorney

state_ratio <- data.frame(StateAbbr = state_data$state,state_data$attorney_client_ratio )

state_ratio

ggplot(state_ratio, aes(x = StateAbbr, y = state_data.attorney_client_ratio)) +
  geom_bar(stat = "identity") +
  labs(x = "State", y = "Ratio") +
  coord_flip()

```
```{r}
final <- data.frame(AverageHours= df_avg$Hours, IncomePerState= income_data$average_state,ClientRatio= state_ratio$state_data.attorney_client_ratio)


attach(final)
plot(AverageHours, IncomePerState, main="Scatterplot Example",
   xlab="Average Hours ", ylab="Client per Attonery ", pch=19)

	
hist(final)

```